Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-07-13 kappa.tex
Add preliminary conclusions.

% ======================================================================
\section{Future perpectives}
% ======================================================================
% Model improvements
% Fieldwork
% Geomorphological-oriented plots

%\begin{figure}
%  \centering
%  \makebox[0pt]{\includegraphics{photo-tuya-moraine}}
%  \caption{The Tuya Lake Moraine in northern British Columbia.}
%  \label{fig:photo-tuya-moraine}
%\end{figure}

>>>>

% ======================================================================
\section{Conclusion and perpectives}
% ======================================================================

In this thesis, I present numerical reconstructions of the former Cordilleran
ice sheet to a level of detail that has, to my knowledge, never been attained
before. The coupled SIA-SSA approximation is crude, yet it is today the only
approach that allows, combined with massively parallel computing, this level of
detail over such time-scales as involved in the growth and decay of ice sheets.

Through sensitivity studies, I have identified the predominant sources of
uncertainty, which I have assessed using a combination of detailed
meteorological datasets and available geological evidence. The simulations
supporting this thesis allowed estimates of the LGM Cordilleran ice sheet
extent and volume, identification of the main ice-dispersal centres, and
partial understanding of the deglaciation patterns. However, there remain much
room for improvement of the model set-up and further analysis of the simulation
output.

% ----------------------------------------------------------------------
\subsection{Precipitation change}
% ----------------------------------------------------------------------

Throughout the thesis, precipitation is prescribed as monthly, spatial grids,
but it is assumed equal to the present values. Although there is no doubt that
the distribution of precipitation has been affected by the evolution of global
climate and the intermittent presence of ice sheets and ice caps, such changes
are difficult to assess. In the North American Cordillera, much of the interior
was likely dryer than today due to the presence of ice caps reinforcing the
topographic barrier formed by the Pacific Coast ranges. However, some regions
may have become wetter as a side effect of redirected winds.

In meteorology, these effects are typically modelled using atmospheric
circulation models. However, I have shown in \CCLI that atmospheric reanalyses,
circulation models using data assimilation techniques, fail to agree on the
present-day pattern of precipitation in the Cordillera due to the highly
mountainous topography, with dramatic effects on the ice sheet model response.
Considering additional difficulties inherent to palaeo-climate modelling, it
can only be expected, that palaeo-climate models without data assimilation will
produce even much higher discrepancies in the ice sheet model response.

% ----------------------------------------------------------------------
\subsection{Geothermal heat flux}
% ----------------------------------------------------------------------

Geothermal heat flux, controlling en-glacial temperature distribution and the
location of basal melt, was assumed constant. However the North America is an
active subduction zone with complex geothermal patterns. Including this
variability in the ice sheet model will offer new opportunities to study its
output.

Most notably, modelled basal flow could be compared with glacial lineations,
one of the main sources of evidence used by geomorphologists to reconstruct
former ice sheets. For instance, in our simulations, large portions of the
Cordilleran ice sheet have remain cold-based throughout the simulation. As
cold-based glaciers typically do not form lineations, this could indicate that
some of the lineation patterns we observe in the North American Cordillera
could predate deglaciation, and instead have been formed during an earlier
phase of advance or retreat of the ice sheet.

% ----------------------------------------------------------------------
\subsection{The Cassiar ice cap}
% ----------------------------------------------------------------------

In the summer of 2012, I have co-organised and participated in fieldwork with
Martin Margold and Ping Fu to collect samples of glacial deposits in the
southern Cassiar Mountains, northern British Columbia. The primary target of
the fieldwork was a triple-crested, sharp, ca. 50\,m-high moraine located in
the vicinity of Tuya Lake the western slope of the Cassiar Mountains
(\cref{fig:photo-tuya-moraine}). Although this is a unique feature in the area,
the size of the moraine, and its location at the foothills of the mountain
range indicate that it was likely formed by the tributary from an ice cap,
rather than a stand-alone valley glacier. Because the relatively low-elevated
Cassiar Mountains do not appear as a major nucleation centre in our simulations
(\CCYC, Fig.~5), it is tempting to think that such a large feature may result
from late-glacial re-advance of a remnant of the decaying ice sheet.
High-resolution modelling of the deglaciation phase, perhaps at a regional
scale, may help to assess this hypothesis.

\begin{figure}
  \centering
  \makebox[0pt]{\includegraphics{photo-tuya-moraine}}
  \caption{The Tuya Lake Moraine in northern British Columbia.}
  \label{fig:photo-tuya-moraine}
\end{figure}
